### 2017-01-30 manuscript.rmd
Update manuscript.rmd
Adds bibliography PCMs

A rich theoretical and statistical methodology of phylogenetic comparative methods have been developed over the last three decades to address the challenges and opportunities of comparisons across species [@Felsenstein:1985ua; @Grafen:1989um; @Chamberlain:2012vx; @Garamszegi:2014ew; @Pagel:1999gc]. These challenges include non-independence of observations across species -- more closely related species will tend to be more similar to each other than distantly related species. Phylogenetic comparative methods have largely been applied to the evolution of morphological and ecological traits, but are just as relevant to CFG. Most CFG studies have abstained from phylogenetic approaches and instead heavily rely on multiple pairwise comparisons (Fig. 1A) between contemporary traits that do not account for the fact that evolutionary relationships explain much of the structure of variation across species. This leaves CFG studies susceptible to serious statistical errors and is a missed opportunity to ask questions that are only accessible in an explicit phylogenetic context.

One reason that CFG has not yet embraced phylogenetic approaches is that it has not yet been concretely demonstrated that pairwise and phylogenetic approaches can lead to different results in CFG studies. In this manuscript we examine this issue. Although the value of phylogenetic approaches has been repeatedly shown in comparative analyses of other types of character data [@Chamberlain:2012vx; @Ricklefs:1996wb], this value is apparently not well known to the functional genomics community since the field has until recently largely focused on within-species analyses. Many initial CFG studies have also been focused on only two species at a time, where the differences between phylogenetic and pairwise approaches are not as pronounced. As the number of species considered in each CFG study grows, the shortcomings of pairwise approaches will become far more acute.

@@

We next examined the possibility that ascertainment biases were differentially impacting the inference of evolution following duplication and speciation events. We focused on two possible sources of bias - node depth and branch length. We found no evidence that either affected our results (Supplementary Materials). We also examined the sensitivity of the results to the calibration times used for speciation events on the gene trees. This is important because it is expected that genes from separate species trace backward in time to a common ancestor at older times than the splitting times of the species to which they belong [Takahata:1989uh, Degnan:2009hr]. There is also uncertainty associated with the timing of these speciation events. In replicate analyses we added random noise to the calibration times and found that it does not impact our inability to reject the null hypothesis (Supplementary Materials).

>>>>

A rich theoretical and statistical methodology of phylogenetic comparative methods have been developed over the last three decades to address the challenges and opportunities of comparisons across species [@Felsenstein:1985ua; @Grafen:1989um; @Pagel:1999gc; Revell:2012jd; @FitzJohn:2012ey; @Uyeda:2014jl; @Garamszegi:2014ew]. These challenges include non-independence of observations across species -- more closely related species will tend to be more similar to each other than distantly related species. Phylogenetic comparative methods have largely been applied to the evolution of morphological and ecological traits, but are just as relevant to CFG. Most CFG studies have abstained from phylogenetic approaches and instead heavily rely on multiple pairwise comparisons (Fig. 1A) between contemporary traits that do not account for the fact that evolutionary relationships explain much of the structure of variation across species. This leaves CFG studies susceptible to serious statistical errors and is a missed opportunity to ask questions that are only accessible in an explicit phylogenetic context.

One reason that CFG has not yet embraced phylogenetic approaches is that it has not yet been concretely demonstrated that pairwise and phylogenetic approaches can lead to different results in CFG studies. In this manuscript we examine this issue. Although the value of phylogenetic approaches has been repeatedly shown in comparative analyses of other types of character data [@Ricklefs:1996wb; @Chamberlain:2012vx; @OMeara:2012gq], this value is apparently not well known to the functional genomics community since the field has until recently largely focused on within-species analyses. Many initial CFG studies have also been focused on only two species at a time, where the differences between phylogenetic and pairwise approaches are not as pronounced. As the number of species considered in each CFG study grows, the shortcomings of pairwise approaches will become far more acute.

@@

We next examined the possibility that ascertainment biases were differentially impacting the inference of evolution following duplication and speciation events. We focused on two possible sources of bias - node depth and branch length. We found no evidence that either affected our results (Supplementary Materials). We also examined the sensitivity of the results to the calibration times used for speciation events on the gene trees. This is important because it is expected that genes from separate species trace backward in time to a common ancestor at older times than the splitting times of the species to which they belong [@Takahata:1989uh; @Degnan:2009hr]. There is also uncertainty associated with the timing of these speciation events. In replicate analyses we added random noise to the calibration times and found that it does not impact our inability to reject the null hypothesis (Supplementary Materials).
